% Emanuel E. Heitlinger - Curriculum Vitae
%

% Set your name here
\def\name{Emanuel G. Heitlinger}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {evolution, ecology, host-parasite interactions,
    transcriptomics, genomics},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}


% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\pagestyle{empty}

% Custom section fonts
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

% Make lists without bullets and compact spacing
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
    \setlength{\itemsep}{0.25em}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0.25em}
  }
}{
  \end{list}
}


% Place name at left
{\huge \name}

% Alternatively, print name centered and bold:
%\centerline{\huge \bf \name}

\bigskip

\begin{minipage}[t]{0.5\textwidth}
  Redtenbacherstr. 9 \\
  76133 Karlsruhe \\
  Germany
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
  Phone: (+49) 721 9822588 \\
  Email: \href{mailto:emanuelheitlinger@gmail.com}{emanuelheitlinger@gmail.com} \\
  Born September, 12th 1980 in Schw\"abisch Gm\"und, Germany\\
\end{minipage}


\section*{Education}

\begin{itemize}
\item [2008-2012] Doctoral studies, Karlsruhe Institute of Technology.
  \begin{itemize}
  \item \textbf{Dissertation:} Divergence of an introduced population
    of the swimbladder-nematode \textit{Anguillicola crassus} - a
    transcriptomic perspective.
  \item \textbf{Supervisors:}
    Prof. Dr. Horst Taraschewski and Prof. Mark Blaxter.
  \end{itemize}
\item [2007-2008] Work on diploma thesis, University of Karlsruhe,
  Zoological Institute, Department for Parasitology and Ecology.
  \begin{itemize}
  \item \textbf{Thesis title:} Vergleichende licht- und
    elektronenmikroskopische Untersuchungen am Intestinaltrakt des
    invasiven Schwimmblasennematoden \textit{Anguillicola crassus} aus
    verschiedenen Aalarten.
  \end{itemize}
\item [2001-2007] Undergraduate studies in Biology, University of
  Karlsruhe.
  \begin{itemize}
  \item \textbf{Main subject:} Zoology
  \item \textbf{Subsidiary subjects:} Genetics, Botany
  \end{itemize}
\item [1991-2000] Secondary school, Privat-Gymnasium St.Paulusheim
  Bruchsal.
\item [1987-1991] Pirmary school, Kraichtal Ober\"owisheim.
\end{itemize}

\section*{Employment}

\begin{itemize}
\item [2008-2011] Research assistant, Karlsruhe Institute of
  Technology, Zoological Institute, Department for Parasitology and
  Ecology.
\item [2000-2001] Alternative military service, youth centre Bruchsal.
\end{itemize}

% \subsection*{Humboldt University Berlin, Department of applied Parasitolgy}
%
% \begin{itemize}
% \item
% \end{itemize}

\section*{Fields of Research Interest}

Ecology and evolution of host-parastite interactions, transcriptomics, genomics

\section*{Research}

\subsection*{Peer Reviewed Publications}

\begin{itemize}
\item Dominik R Laetsch, \textbf{Emanuel G Heitlinger}, Horst
  Taraschewski, Steven A Nadler and Mark L Blaxter (2012) The
  phylogenetics of Anguillicolidae (Nematoda: Anguillicolidea),
  swimbladder parasites of eels. \textit{under review} \textit{BMC
    Evolutionary Biology}.
\item \textbf{Emanuel G Heitlinger}, Dominik R Laetsch, Urszula
  Weclawski, Yu-San Han and Horst Taraschewski (2009)
  \href{http://www.parasitesandvectors.com/content/2/1/48}{Massive
    encapsulation of larval \textit{Anguillicoloides crassus} in the
    intestinal wall of Japanese eels}. \textit{Parasites \& Vectors},
  2:48.
\end{itemize}

% \subsection*{Scientific Software}
%
% \begin{itemize}
% \item \href{http://foo}{foo},
%   A R package
% \end{itemize}

% \subsection*{Other Publications}

% \begin{itemize}
% \item A book
% \end{itemize}

% \section*{Teaching}

% \subsection*{Humboldt University Berlin}

% \begin{itemize}
% \item
% \end{itemize}

\subsection*{Conference Presentations}

\begin{itemize}

\item 3rd Status Symposium, Volkswagen Foundation Funding Initiative
  Evolutionary Biology, November 7-11 2011, Sylt, Germany. Oral
  presentation: ``Divergence of an intoduced parasite: a
  transcriptomic perspective on \textit{Anguillicola crassus}''.

\item 2nd Status Symposium, Volkswagen Foundation Funding Initiative
  Evolutionary Biology, May 9-12 2010, Frauenchiemsee, Germany. Oral
  presentation: ``The transcriptome of \textit{Anguillicoloides
    crassus} sampled by pyrosequencing''.

\item 24th Biannual conference of the German society of parasitology
  (DGP), March 16-19 2010, M\"unster, Germany. Two oral presentations:
  ``The transcriptome of \textit{Anguillicoloides crassus} sampled by
  pyrosequencing'' and ``Massive encapsulation of larval
  \textit{Anguillicoloides crassus} in the intestinal wall of the
  Japanese eel''.

\item Mind the gap: joining empirical and theoretical population
  genetics, October 2-3 2009, Freiburg, Germany. Oral Presentation:
  ``Divergence between European and Asian populations of the
  swimbladder nematode \textit{Anguillicoloides crassus}''.

\item 1st Status Symposium, Volkswagen Foundation Funding Initiative
  Evolutionary Biology, February 25-27 2009, M\"unster,
  Germany. Poster: ``Divergence between East Asian and European
  populations of the swimbladder-nematode \textit{Anguillicola
    crassus}''.

\item Xth European Multicolloquium of Parasitology - EMOP 10, August
  24-28, 2008, Paris, France. Oral Presentation: ``Divergence between
  East Asian and European populations of the swimbladder-nematode
  \textit{Anguillicola crassus}''.

\end{itemize}

% \section*{Professional Activities}

% \begin{itemize}
% \item Reviewer for
%   \textit{foo}
% \item Departmental service:
%   Faculty Recruiting Committee (2010-11, 2
%   G.S. Maddala Prize Committee (2010-11).
% \item Member, foo, 2006--Present.
% \item Member, foo2, 2010--Present.
% \end{itemize}

% \section*{Advising}

% \begin{itemize}
% \item Ph.D. Students (secondary advisor):
%   foo foo (Ph.D. candidate), % candidacy exam, 2011
%   Foo  (Ph.D. candidate),          % candidacy exam, 2011

% \item Undergraduate Students:
%   foo (2011).
% \end{itemize}

\section*{Honors, Awards, \& Fellowships}

\begin{itemize}
\item [2008] Volkswagen Stiftung PhD Fellowship, Funding Initiative
  Evolutionary Biology, full funding of research position and research
  material
\end{itemize}

% \section*{Miscellaneous}

% \begin{itemize}
% \item \textit{Security Clearance:} Special agent.
% \end{itemize}

% Footer
\begin{center}
  \begin{small}
    Last updated: \today
  \end{small}
\end{center}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End: 
