
% this file is called up by thesis.tex
% content in this file will be fed into the main document

\chapter{Materials \& methods} % top level followed by section, subsection


% ----------------------- paths to graphics ------------------------

% change according to folder and file names
\ifpdf
    \graphicspath{{8/figures/PNG/}{8/figures/PDF/}{8/figures/}}
\else
    \graphicspath{{8/figures/EPS/}{8/figures/}}
\fi

% ----------------------- contents from here ------------------------


\section{Sampling of worms from wild eels for Sanger- and
  pyrosequencing}

Cultured eels were acquired from an aquaculture directly adjacent to
Kaoping river (22.6418N; 120.4440E) 15km stream upwards from its
estuary. Wild eel were bought from a fisherman, fishing in the estuary
of Kao-Ping river (22.5074N; 120.4220E). All eels were transported to
the Institute of Fisheries Science at the National Taiwan University
in Taipei in aerated plastic bags, where they were sheltered until
dissection.

Eels were decapitated, length (to the nearest 1.0mm) and weight (to
the nearest 0.1g) were measured, and sex was determined by visual
inspection of the gonads. The swimbladder was opened, adult worms were
removed from the lumen with a forceps, their sex was determined, and
they were counted. All adult \textit{A. crassus} were preserved in
RNAlater(Quiagen, Hilden, Germany) in individual plastic tubes.

Worms from the European eel were sampled in Sniardwy Lake, Poland
(53.751959N; 21.730957E) by Urszula Weclawski and from the
Linkenheimer Altrhein, Germany (49.0262N; 8.310556E), following a
procedure similar to the one described above for worms from Taiwan.

\section{RNA-extraction and cDNA synthesis for Sanger- and
  pyrosequencing}

Total RNA was extracted from single, whole worms using the RNeasy kit
(Quiagen, Hilden, Germany), following the manufacturers
protocol. Alternatively parts of the liver of the host species
\textit{Anguilla japonica}, which also had been preserved in RNAlater
were used for RNA extraction, following the same protocol.

The Evrogen MINT cDNA synthesis kit (Evrogen, Moscow, Russia ) was
then used to amplify mRNA transcripts according to the manufacturers
protocol. It uses an adapter sequence at 3' the end of a poly
dT-primer for first strand synthesis and adds a second adapter
complementary to the bases at the 5' end of the transcripts by
terminal transferase activity and template switching. Using these
adapters it is possible to specifically amplify mRNA enriched for
full-length transcripts.

\section{Cloning for Sanger-sequencing}

The obtained cDNA preparations were undirectionally cloned into
TOPO2PCR-vectors (Invitrogen, Carlsbad, USA) and TOP10 chemically
competent cells (Invitrogen, Carlsbad, USA) were transformed with this
construct. The cells were plated on LB-medium-agarose containing
Kanamycin (5mg/ml), xGal
(5-bromo-4-chloro-3-indolyl-$\beta$-D-galactopyranoside) and IPTG
(Isopropyl-$\beta$-D-1-thiogalactopyranosid). After 24h of incubation
at $36\,^{\circ}\mathrm{C} $ cells were picked into 96-well
micro-liter-plates containing liquid LB-medium and Kanamycin (5mg/ml)
and incubated for another 24h. Subsequently 2ml of the cells were used
as template for amplification of the insert by PCR using the primers
\begin{description}
\item[Forward] M13F(GTAAAACGACGGCCAGT) and
\item[Reverse] M13R(GGCAGGAAACAGCTATGACC)
\end{description}
in a concentration of 10$\mu$M. The protocol for PCR cycling is shown
% in table \ref{tab:PCR}.

\begin{table}[h]
  \centering
  \begin{tabular}{lllll} 
    \textbf{Initial denaturation} &  $ 94\, ^{\circ}\mathrm{C} $ & 5min &  &\\ 
    \hline
    \textbf{Denaturation} &  $ 94\, ^{\circ}\mathrm{C} $ &30s& & \\ 
    \textbf{Annealing} &   $ 54\, ^{\circ}\mathrm{C} $ & 45s & 35 cycles &\\ 
    \textbf{Elongation} &   $ 72\, ^{\circ}\mathrm{C} $ & 2min &  &\\ 
    \hline
    \textbf{Final Elongation} &   $ 72\, ^{\circ}\mathrm{C} $ & 10min &\\ 
  \end{tabular}   
  \caption{PCR protocol for insert amplification}
  \label{tab:PCR}
\end{table}

Amplification products were controlled on gel and cleaned using SAP
(Shrimp Alkaline Phosphatase) and ExoI (Exonuclease I). Sequencing
reactions were performed using the BigDye-Terminator kit and
PCR-primers (forward or reverse) in a concentration of 3.5$\mu$M and
sequenced on an ABI 3730 DNA Analyzer (Applied Biosystems, Foster
City, California, USA).  For \textit{A. crassus} the following
libraries were prepared:
 
\begin{description}
\item{Ac\_197F:} Female from Taiwanese aquaculture
\item{Ac\_106F:} Female from Taiwanese aquaculture
\item{Ac\_M175:} Male from Taiwanese aquaculture
\item{Ac\_FM:} Female from Taiwanese aquaculture
\item{Ac\_EH1:} Same cDNA preparation as Ac\_FM, but sequenced by
  students in a practical
\end{description}

For \textit{Anguilla japonica} the following three libraries:
\begin{description}
\item{Aj\_Li1:} liver of an eel from aquaculture
\item{Aj\_Li2:} liver of an eel from aquaculture
\item{Aj\_Li3:} liver of an eel from aquaculture
\end{description}

\section{Pilot Sanger-sequencing}

The original sequencing-chromatographs ("trace-files") were renamed
according to the NERC environmental genomics scheme. "Ac" was used as
project-identifier for \textit{Anguillicola crassus}, "Aj" for
\textit{An. japonica}. In \textit{Anguillicola} sequences information
on the sequencing primer (forward or reverse PCR primer;
\textit{An. japonica} sequences were all sequenced using the forward
PCR primer) was stored in the middle "library"-field, resulting in
names of the following form:

\begin{itemize}
\item \texttt{Ac\_[\textbackslash{}d|\textbackslash{}w]\{2,4\}(f|r)\_\textbackslash{}d\textbackslash{}d\textbackslash{}w\textbackslash{}d\textbackslash{}d}
\item  \texttt{Aj\_[\textbackslash{}d|\textbackslash{}w]\{2,4\}\_\textbackslash{}d\textbackslash{}d\textbackslash{}w\textbackslash{}d\textbackslash{}d}
\end{itemize}

The last field indicates the plate number (two digits), the row (one
letter) and the column (two digits) of the corresponding clone. For
first quality trimming \texttt{trace2seq}, a tool derived from
\texttt{trace2dbEST} (both part of \texttt{PartiGene}
\cite{parkinson_partigene--constructing_2004}) was used, briefly it
performs quality trimming using \texttt{phred}
\cite{ewing_base-calling_1998} and trimming of vector sequences using
\texttt{cross-match} \cite{PHRAP}. The adapters used by the MINT kit
were trimmed by supplying them in the vector-file used for trimming
along with the TOPO2PCR-vector. After processing with
\texttt{trace2seq} additional quality trimming was performed on the
produced sequence-files using a custom script. This trimming was
intended to remove artificial sequences produced when the sequencing
reaction starts at the 3' end of the transcript at the poly-A
tail. These sequences typically consist of numerous homo-polymer-runs
throughout their length caused by "slippage" of the reaction.
The basic Perl regular expression used for this was:\\

\texttt{/(.*A\{5,\}|T\{5,\}|G\{5,\}|C\{5,\}.*)\{\$lengthfac,\}/g}\\

Where \texttt{\$lengthfac} was set to the length of the sequence
divided by 70 and rounded to the next integer. So only one
homo-polymer-run of more then 5 bases was allowed per 75 bases.

Sequences were screened for host contamination by a comparison of
\texttt{Blast} searches against nempep \cite{pmid21550347} (version 4)
and a fish protein database. Sequences producing better bit scores
against fish proteins than nematode proteins were labeled as
host-contamination.

Only the trace-files corresponding to the sequences still regarded as
good after this step were processed with
\texttt{trace2dbEST}. Additionally to the processing of traces already
included in \texttt{trace2seq} sequences were preliminary annotated
using \texttt{Blast} versus the NCBI-NR non-redundant protein database
and EST-submission-files were produced.

\section{Pyrosequencing}

\subsection{cDNA preparation and sequencing}

RNA was extracted from individual adult male and female nematodes and
from a population of L2 larvae. RNA was reverse transcribed and
amplified into cDNA using the MINT-cDNA synthesis kit (Evrogen,
Moscow, Russia).  For host contamination screening a liver-sample from
an uninfected \textit{An. japonica} was also processed. Emulsion PCR
was performed for each cDNA library according to the manufacturer's
protocols (Roche/454 Life Sciences), and sequenced on a Roche 454
Genome Sequencer FLX. All samples were sequenced using the FLX
Titanium chemistry, except for the Taiwanese female sample T2, which
was sequenced using FLX standard chemistry, to generate between 99,000
and 209,000 raw reads. For the L2 larval library, which had a larger
number of non-\textit{A. crassus}, non-\textit{Anguilla} reads,
screening Roche 454 data produced on the same run in independent
sequencing lanes confirmed that these data were not laboratory
contaminants.

\subsection{Trimming, quality control and assembly}

Raw sequences were extracted in \texttt{fasta}-format (with the
corresponding qualities files) using \texttt{sffinfo} (Roche/454) and
screened for adapter sequences of the MINT-amplification-kit using
\texttt{cross-match} \cite{PHRAP} (with parameters \texttt{-minscore
  20 -minmatch 10}). \texttt{Seqclean} \cite{tgicl_pertea} was used to
identify and remove poly-A-tails, low quality, repetitive and short
(<100 base) sequences. All reads were compared to a set of screening
databases using \texttt{Blast} (expect value cutoff E<1e-5, low
complexity filtering turned off: -F F). The databases used were (a) a
host sequence database comprising an assembly of the
\textit{An. japonica} Roche 454 data, an unpublished assembly of
\textit{An. anguilla} Sanger dideoxy sequenced expressed sequence tags
(made available to us by Gordon Cramb, University of St Andrews) and
transcripts from EeelBase \cite{pmid21080939} a publicly available
transcriptome database for the European eel; (b) a database of
ribosomal RNA (rRNA) sequences from eel species derived from my Roche
454 data and EMBL-Bank; and (c) a database of rRNA sequences
identified in my \textit{A. crassus} data by comparing the reads to
known nematode rRNAs from EMBL-Bank. This last database notably also
contained xenobiont rRNA sequences. Reads with matches to one of these
databases over more than 80\% of their length and with greater than
95\% identity were removed from the dataset. Screening and trimming
information was written back into sff-format using \texttt{sfffile}
(Roche 454). The filtered and trimmed data were assembled using the
combined assembly approach \cite{pmid20950480}: two assemblies were
generated, one using \texttt{Newbler v2.6} \cite{pmid16056220} (with
parameters \texttt{-cdna -urt}), the other using \texttt{Mira v3.2.1}
\cite{miraEST} (with parameters
\texttt{--job=denovo,est,accurate,454}). The resulting two assemblies
were combined into one using \texttt{Cap3} \cite{Cap3_Huang} at
default settings and contigs were labeled by whether they derived from
both assemblies or one assembly only.

\subsection{Evaluation of the assemblies}

The ace-files for all three (two first-order, one second-order)
assemblies were interrogated for the fate of single reads. This was
used to tabulate the full read-first-order-second-order-associations.

\texttt{Blast} (\texttt{blastx -e 1e-5}) was used to search the
complete proteomes of \textit{C. elegans} (as present in wormbase
v.220) and the complete proteome of \textit{B. malayi} (as present in
uniref 100) for the contigs and singletons of all investigated
assemblies. A custom Perl-script (provided by S. Kumar) was then used
to mask all bases in the database covered. For each sequence in the
database the size of the masked region was then determined and
statistics were created summarising the number of database-sequences
with any coverage, the number with coverage over 80\% of their
sequence-length and the overall proportion of bases covered.

Based on reads shared between clusters I collapsed reads linked by
such read-paths, assigned a cluster-id and recorded the size of the
cluster.

To estimate contig-coverage I converted \texttt{sam}-output generated
with \texttt{ssaha2}\cite{pmid11591649} via a sorted \texttt{bam}-file
to \texttt{pileup}-format using \texttt{samtools}
\cite{journals/bioinformatics/LiHWFRHMAD09}. For a second evaluation
I excluded best-hits mapping to multiple contigs before converting the
\texttt{sam}-file to obtain unique coverage.

\subsection{Post-assembly classification and taxonomic assignment of
  contigs}

After assembly contigs were assessed a second time for host and other
contamination by comparing them (using \texttt{Blast}) to the three
databases defined above, and also to nembase4, a nematode
transcriptome database derived from whole genome sequencing and EST
assemblies \cite{parkinson_nembase:resource_2004, pmid21550347}. For
each contig, the highest-scoring match was recorded as long as it
spanned more than 50\% of the contig. I also compared the contigs to
the NCBI non-redundant nucleotide (NCBI-nt) and protein (NCBI-nr)
databases, recording the taxonomy of all best matches with expect
values better than 1e-05. TUGs with a best hit to non-Metazoans and to
Chordata within Metazoa were additionally excluded from further
analysis.

\subsection{Protein prediction and annotation}
\label{pro4est}

Protein translations were predicted from the contigs using
\texttt{prot4EST} (version 3.0b)
\cite{wasmuth_prot4est:_2004}. Proteins were predicted either by
joining single high scoring segment pairs (HSPs) from a \texttt{Blast}
search of uniref100 \cite{pmid18836194}, or by \texttt{ESTscan}
\cite{estscan}, using as training data the \textit{Brugia malayi}
complete proteome back-translated using a codon usage table derived
from the \texttt{Blast} HSPs, or, if the first two methods failed,
simply the longest ORF in the contig. For contigs where the protein
prediction required insertion or deletion of bases in the original
sequence, I also imputed an edited sequence for each affected
contig. Annotations with Gene Ontology (GO), Enzyme Commission (EC)
and Kyoto Encyclopedia of Genes and Genomes (KEGG) terms were inferred
for these proteins using \texttt{Annot8r} (version 1.1.1)
\cite{schmid_annot8r:_2008}, using the annotated sequences available
in uniref100 \cite{pmid18836194}. Up to 10 annotations based on a
\texttt{Blast} similarity bitscore cut-off of 55 were obtained for
each annotation set. The complete \textit{B. malayi} proteome (as
present in uniref100) and the complete \textit{C. elegans} proteome
(as present in wormbase v.220) were also annotated in the same
way. \texttt{SignalP V4.0} \cite{pmid21959131} was used to predict
signal peptide cleavage sites and signal anchor signatures for the
\textit{A. crassus}-transcriptome and similarly again for the
proteomes of the tow model-worms.  Additionally \texttt{InterProScan}
\cite{pmid11590104} (command line utility \texttt{iprscan} (version
4.6) with options \texttt{-cli -format raw -iprlookup -seqtype p
  -goterms}) was used to obtain domain based annotations for the high
credibility assembly (highCA) derived contigs.

I recorded the presence of a lethal RNAi-phenotype in the
\textit{C. elegans} ortholog of each TUG using the biomart-interface
\cite{pmid22083790} to wormbase v. 220 through the R-package
\texttt{biomaRt} \cite{pmid19617889}.

\subsection{Single nucleotide polymorphism analysis}

I mapped the raw reads against the the complete set of contigs,
replacing imputed sequences for originals where relevant, using
\texttt{ssaha2} \cite{pmid11591649} (with parameters \texttt{-kmer 13
  -skip 3 -seeds 6 -score 100 -cmatch 10 -ckmer 6 -output sam -best
  1}). From the \texttt{ssaha2} output, pileup-files were produced
using \texttt{samtools} \cite{journals/bioinformatics/LiHWFRHMAD09},
discarding reads mapping to multiple regions. \texttt{VarScan}
\cite{pmid19542151} (\texttt{pileup2snp}) was used with default
parameters on pileup-files to output lists of single nucleotide
polymorphisms (SNPs) and their locations. For enrichment analysis of
GO-terms I used the R-package \texttt{GOstats} \cite{pmid17098774}.

Using \texttt{Samtools} \cite{journals/bioinformatics/LiHWFRHMAD09}
(\texttt{mpileup -u}) and \texttt{Vcftools} \cite{pmid21653522}
(\texttt{view -gcv}) I genotyped individual libraries for the list of
previously found overall SNPs. Genotype-calls were accepted at a
phred-scaled genotype quality threshold of 10. In addition to the
relative heterozygosity (number of homozygous sites/number of
heterozygous sites) I used the R package \texttt{Rhh}
\cite{pmid21565077} to calculate internal relatedness
\cite{pmid11571049}, homozygosity by loci \cite{pmid17107491} and
standardised heterozygosity \cite{coltman81j} from these data.

Using 1000 bootstrap replicates I confirmed the significance of
heterozygote-heterozygote correlation by analysing the mean and 95\%
confidence intervals from 1000 bootstrap replicates estimated for all
measurements.

\subsection{Gene-expression analysis}

Read-counts were obtained from the \texttt{bam}-files generated also
for genotyping using the R-package \texttt{Rsamtoools}
\cite{rsamtools}. TUGs with less than 48 reads over all libraries were
excluded from analysis, as diagnostic plot (not shown) indicated a
lack of statistical power for lower overall expression. I used the
R-package \texttt{DESeq} \cite{pmid20979621} (version 1.6.1) to assess
statistical significance of differences in counts according to groups
of libraries.

Additionally I collapsed TUGs by their orthologous assignment in
\textit{C. elegans} and \textit{B. malayi}. I used the sums of counts
for these orthologous-groups to asses the influence of mapping to my
potentially fragmented reference. For both model-nematodes fold-change
and p-values were obtained the same way than for the contigs and
merged with these.

\subsection{Enrichment analysis}
\label{GO-over-meth}

We used Mann-Whitney u-tests to test the influence of factors on dn/ds
values, when multiple contrasts between groups (factors) were
investigated we used Nemenyi-Damico-Wolfe-Dunn tests. For
overrepresentation of one group (factor) in another group (factor) we
used Fisher's exact test.

Prior to analysis of GO-term over-representation (based on dn/ds or
expression values) we used the R-package \texttt{annotationDbi}
\cite{AnnotationDbi} to obtain a full list of associations (also with
higher-level terms) from \texttt{annot8r}-annotations. We then used the
R-package \texttt{topGO} \cite{topGO} to traverse the annotation-graph
and analyse each node in the annotation for over-representation of the
associated term in the focal gene-set compared to an appropriate
universal gene-set (all contigs with dn/ds values or all contigs
analysed for gene-expression) with the ``classic'' method and Fisher's
exact test.

\section{Transcriptomic divergence in a common garden experiment}

\subsection{Experimental infection of eels}

\textit{An. anguilla} were obtained from the Albe-Fishfarm in
Haren-R\"utenbrock, Germany. \textit{An. japonica} were caught at the
glass-eel stage in the estuary of Kao-ping River, Taiwan by
professional fishermen and kept at a water temperature of
26$^{\circ}$C until they reached a size of $>$ 35 cm.

The absence of infections with \textit{A. crassus} in both eel-species
was confirmed by dissection of 10 individuals of each species.

After an acclimatisation period of 4 weeks (\textit{An. anguilla}) or
when they reached a size of $>$ 35cm (\textit{An. japonica}) eels were
infected using a stomach tube as described in
\cite{boon1990effect}. During the infection period water temperature
was held constant at 20$^{\circ}$C. Eels were kept in 160-litre tanks
in groups of 5-10 individuals and continuously provided with fresh,
oxygenated water and once every two days with commercial fish pellets
(Dan-Ex 2848, Dana Feed A/S Ltd, Horsens, Denmark) \textit{at
  libitum}.

L2 larvae used for the infection were collected from the swimbladders
of wild yellow and silver eels from the River Rhine near Karlsruhe and
from Lake M\"uggelsee near Berlin in Germany. Taiwanese larvae were
obtained from eels from an aquaculture adjacent to Kao Ping River in
south Taiwan and from a second aquaculture in Yunlin county,
approximately 150 km further north on the west coast of Taiwan. They
were stored at 4$^{\circ}$C for no longer than 2 weeks before copepods
were infected. Mixed species samples of uninfected copepods were
collected from a small pond near Karlsruhe, known to be free of eels
(and \textit{Anguillicola}). They were infected individually in wells
of micro-titer plates at an intensity of roughly 10 L2-larvae per
copepod. One week after infection they were placed in bigger
tanks. Twice a week yeast was provided as food and at 21 dpi infective
L3 were harvested with using a tissue potter as described by
\cite{haenen_improved_1994}. 50 L3 for infection of individual eel
were suspended in 100 $\mu$l RPMI-1640 medium (Quiagen, Hilden,
Germany) and eels were infected as described above.

55-57 days post infection (dpi) eels were euthanized and dissected.
The swimbladder was opened and after determination of the sex of adult
worms under a binocular microscope (Semi 2000, Zeiss, Germany), adult
\textit{A. crassus} were immediately immersed in RNAlater (Quiagen,
Hilden, Germany).

\subsection{RNA extraction and preparation of sequencing libraries}

RNA was extracted from 12 individual female worms and for 12 pools of
male worms using the RNeasy-kit (Quiagen, Hilden, Germany) (see table
\ref{tab:lib-prep}).

The paired-end TruSeqTM RNA sample preparation kit (Illumina) was
followed to build sequencing libraries with insert sizes of roughly
270 bp for paired-end sequencing from cDNA libraries: briefly, poly-T
oligo-attached magnetic beads were used for purification of mRNA and
to simultaneously fragment the RNA. The RNA was then primed with
random hexamer primers for cDNA synthesis and reverse transcribed into
first strand cDNA using reverse transcriptase. The cDNA was cleaned
from the second strand reaction, overhangs were repaired to form blunt
ends, a single ``A''-nucleotide was added at the 3' end and paired end
sequencing adapters ware ligated with a complementary
``T''-overhang. In this step multiple differently indexed paired-end
adapters were used to enable multiplexing of the 24 different
sequencing libraries in 3 pools of 8 samples each. These three pools
all contained one random replicate each for each treatment combination
ensuring complete statistical independence of replicates. Molecules
having adapter sequences were enriched in the mix using PCR and the
libraries were controled for quality and quantity on the BioAnalyzer
(Agilent). Clusters were generated by bridge amplification. The
resulting clusters were sequenced on the Genome Analyzer IIX in
combination with the paired-end module. The first read was sequenced
using using the first primer Rd1 SP. The original template strand was
then used to regenerate the complementary strand, the original strand
was removed and complementary strand acted as a template for the
second read, sequenced primed by the second sequencing primer Rd2 SP.

\subsection{Mapping and normalisation of read-counts}

All sequencing reads were mapped to the fullest 454 assembly (as
defined in \ref{sec:final-full-assembly}; I were including TUGs
inferred as host or xenobiont origin as filter) using \texttt{BWA}
\cite{pmid20080505} (version 0.5.9-r16; \texttt{BWA aln} and
\texttt{BWA sampe} with default options) and processed with
\texttt{samtools} \cite{journals/bioinformatics/LiHWFRHMAD09} (version
0.1.18; \texttt{samtools view -uS -q 1}) to only allow uniquely
mapping reads. All reads mapping to host- and xenobiont off-target
data were removed during downstream evaluation.

Counts were summed for technical replicates and counts to
lowCA-derived contigs were disregarded for statistics on a
contigs-base as well as spurious read counts to contigs with less
than 32 mapping reads in total (see however \ref{collapse} for how
these counts were used in further tests of reference fragmentation).

The remaining counts were normalised using \texttt{DESeq} (version
1.6.1) (i.e.  the normalisation factor was estimated by the median of
scaled counts, similar to the weighted trimmed mean of the log
expression ratios used later in \texttt{edgeR}). All tables
summarising read-counts are based on these normalised counts. I
obtained ``variance stabilised data'' in an expression matrix for each
gene and library using the ``blind'' option in a calculation not
informed (and biased by) the model-design. These data were used in all
gene-centring heatmap and multivariate visualisations. Additionally this
matrix was transposed to get sample-to-sample distances.

\subsection{Statistical analysis with generalised linear models (GLMs)}

The R-package \texttt{edgeR} (version 2.4.1) \cite{pmid19910308} was
used to build negative binomial generalised linear models, as these
specialised GLMs outperformed GLMs in \texttt{DESeq} in speed and
reliability of convergence. Modeled were based on a negative binomial
distribution and the dispersion parameter for each transcript was
approximated with a trend depending on the overall level of
expression. In the maximal fitted model expression was regressed on
worm-sex, host-species and parasite population, including all their
interactions. The full model thus contained terms $ S_i + H_j + P_k +
(SH)_{ij} + (SP)_{ik} + (HP)_{jk} + (SHP)_{ijk} + \varepsilon$, where
$\varepsilon$ is the residual variance, $S_i$ is the effect of the ith
sex (male or female), $H_j$ is the effect of the ith host species
(\textit{An. anguilla} or \textit{An. japonica}), $P_k$ is the effect
of the kth population (European or Asian), $(SH)_{ij}$ is the
sex-by-species interaction and similarly for the other interactions.

The hierarchical nature of generalised linear models was respected
considering (removing) all interaction effects of a main-term
(e.g. $(SP)_{ik}$, $(SH)_{ij}$ and $(SHP)_{ijk}$) when analysing
models for the significance of that term (e.g. $S_i$). Resulting
p-values were corrected for multiple testing using the method of
Benjamini and Hochberg \cite{benjamini1995controlling} and
differential expression was inferred at a false discovery rate (FDR)
of 5\% (adjusted p-value of 0.05).

\subsection{Count-collapsing for orthologs from two model-species}
\label{collapse}

In order to test the influence of deficiencies (i.e. fragmentation) of
the assembly on mapping and read-counts I summed read counts over
orthologous sequence in \textit{C. elegans} and
\textit{B. malayi}. For this purpose I used all reference contigs
(also lowCA-derived contigs to allow inference of fragmentary mapping
to those, but not contigs of non-\textit{A. crassus} origin).
Differential expression for these orthologous-counts was analysed the
same way as for contigs. Contigs were filtered based on inference
from orthologous counts merging the two orthologous evaluations and
the contig evaluation. Differential expression was accepted at a FDR
of 5\% for the contig evaluation and 10\% for both of the two
orthologous evaluations.

\subsection{Multivariate confirmation of linear models}
\label{sec:mult-conf-line}

I used the R-package \texttt{vegan} (version 2.0-2) to perform
constrained redundancy analysis on contigs identified as significant
in GLMs before. For each set of contigs (different for sex, eel-host
or worm-population) the appropriate constrained component was
used. The proportion of the variance explained by the constrained
component was recorded and the constrained component was tested for
significance using a permutation test implemented in \texttt{vegan}.

\subsection{GO-term enrichment analysis}

Enrichment analysis was performed as described above for
pyrosequencing data (see \ref{GO-over-meth}).

\subsection{Clustering analysis}

The R-package \texttt{HeatmapPlus} was used on variance stabilised
expression values to visualise hierarchical clusters similar to the
method of \cite{pmid9843981}. The results were displayed along with
annotations stored in a Bioconductor eSet-class object.

\section{General coding methods}

The bulk of analysis (unless otherwise cited) presented in this paper
was carried out in \texttt{R} \cite{R_project} using custom scripts.
I used a method provided in the R-packages \texttt{Sweave}
\cite{lmucs-papers:Leisch:2002} and \texttt{Weaver} \cite{weaver} for
``reproducible research'' combining \texttt{R} and \LaTeX code in a
single file. The complete reproducible compilations were only carried
out for sub-chapters of this document, the thesis-document was then
compiled from plain \LaTeX sub-documents. Nevertheless all
intermediate data files needed to compile sub-document of the thesis
from data-sources are provided upon request. For general visualisation
I used the R-packages \texttt{ggplot2} \cite{ggplot-book} and
\texttt{VennDiagram} \cite{pmid21269502}.


%%% Local Variables: ***
%%% mode:latex ***
%%% TeX-master: "../thesis.tex"  ***
%%% tex-main-file: "../thesis.tex" ***
%%% End: ***
 






