% this file is called up by thesis.tex
% content in this file will be fed into the main document

% Glossary entries are defined with the command \nomenclature{1}{2}
% 1 = Entry name, e.g. abbreviation; 2 = Explanation
% You can place all explanations in this separate file or declare them in the middle of the text. Either way they will be collected in the glossary.

% required to print nomenclature name to page header
\markboth{\MakeUppercase{\nomname}}{\MakeUppercase{\nomname}}


% ----------------------- contents from here ------------------------

%  
\nomenclature{SNP}{Single Nucleotide Polymorphism; variation occurring
in a single nucleotide between two closely related homlogous
sequences. Leading for example to allelic differences within a
population - or even between homologous chromosomes in an individual.}

\nomenclature{SAGE}{Serial Analysis of Gene Expression; analysis of
gene-expression by sequencing of short, ligated sequence-tags made
from cDNA.}

\nomenclature{ORF}{Open Reading Frame; a region in a DNA-sequence
begining with a start-codon and not containing a stop-condon. For
example a region within a processed mRNA transcript being transcribed
into a protein.}

\nomenclature{dpi}{Days post infection; In infection experiments, a
point in time given in days after an individual has been infected.}

\nomenclature{DNA}{Desoxy Ribonucleic Acid; a chemical molecule
  bearing the heritable genetic information in all life on earth.}

\nomenclature{cDNA}{complementary DNA; from RNA reverse transcribed
  DNA. This is achieved using RNA-dependent DNA polymerase polymerase.}

\nomenclature{PCR}{Polymerase Chain Reaction; a technique for the
amplification of DNA by cyles of DNA-melting, annealing of primers and
enzymatic replication. Heat-stable DNA-polymerase allows the use of
thermal-cycling.}

\nomenclature{RNA}{Ribonucleic Acid; }

\nomenclature{mRNA}{Messenger RNA; }

\nomenclature{NGS}{Next Generation Sequencing; }

\nomenclature{EST}{Expressed Sequence Tag; randomly sampled sequence
frome expressed RNA using Sequencing of cDNA preparations}